MSSM-Penn-Pitt - ACC and DLPFC - FEATURECOUNTS - GRCh38 - CQN - Iterative Model

### 2020-05-04 code/CMC_ExpVal_DataProcessing.Rmd
Validation

# Primary variable of interest
postAdjustCovars = c('Institution', 'Reported_Gender') #c('Institution', 'Reported_Gender'); #'FlowcellBatch'
#postAdjustCovars = c('Institution', 'Reported_Gender');
# Assign residual covariates
residualCovars = setdiff(preAdjustedSigCovars$significantCovars, c(postAdjustCovars, primaryVariable, 'Individual_ID'))
residualSigCovars = preAdjustedSigCovars
covariatesEffects = preAdjustedSigCovars$Effects.significantCovars[residualCovars]
postAdjustCovars = c(postAdjustCovars, names(which.max(covariatesEffects))) %>% unique()
notEstimable = c()

@@

  ggsave( filename=paste0( Tissue, "_Val.pdf" ),
    plot = ggplot( data = Correlation, aes( x = Correlation, y = Log10_Pval, col = Sig ) ) + 
      geom_point() + ggtitle("Log10( PVal) vs. Correlation" ) + 
      theme(plot.title = element_text(hjust = 0.5)) + ylab("-Log( PValue)") ,
    device = "pdf",
    path = NULL,
    scale = 1,
    width = 21,
    height = 7,
    units = "in",
    dpi = 300,
    limitsize = TRUE
  )

@@

@@

#Set Used SynIDs For Provenance
#Syns_Used <- c("")

@@

#TODO add provenance used = Syns_Used,

# Store SVA results
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'CMC_SVA_Diagnosis_Corrected.tsv', sep = '\t', quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='CMC_SVA_Diagnosis_Corrected.tsv', name = 'Normalised, covariates, surrogate and Diagnosis variable adjusted residual expression (for eQTL)', parentId=CODE$properties$id ), activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'residualGeneExpForeQTL'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store residual gene expression for network analysis
RESIDUAL.NET.GENE_EXPRESSION %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'CMC_Diagnosis_Regressed_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='CMC_Diagnosis_Regressed_netResidualExpression.tsv', name = 'Normalised, covariates and Diagnosis removed residual expression (for network analysis)', parentId=CODE$properties$id ), activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$dataSubType = 'residualGeneExpForNetAnlz'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

#Upload Plots:
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ACC_Val.pdf', name = 'LogP value as a function of correlation for ACC', parentId=CODE$properties$id ), activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$tissueTypeAbrv = 'ACC'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ACC_Cor_All.eps', name = 'Histogram of all correlation values for ACC', parentId=CODE$properties$id ), activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$tissueTypeAbrv = 'ACC'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ACC_Cor_gt5Perc.eps', name = 'Histogram of Significant correlation values for ACC', parentId=CODE$properties$id ), activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$tissueTypeAbrv = 'ACC'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='PFC_Val.pdf', name = 'LogP value as a function of correlation for DLPFC', parentId=CODE$properties$id ), activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$tissueTypeAbrv = 'DLPFC'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='PFC_Cor_All.eps', name = 'Histogram of all correlation values for DLPFC', parentId=CODE$properties$id ), activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$tissueTypeAbrv = 'DLPFC'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='PFC_Cor_gt5Perc.eps', name = 'Histogram of Significant correlation values for DLPFC', parentId=CODE$properties$id ), activityName = activityName, executed = thisFile, activityDescription = activityDescription)
  all.annotations$tissueTypeAbrv = 'DLPFC'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

>>>>

# Primary variable of interest
postAdjustCovars = c('Institution', 'Reported_Gender', 'Dx' ) #c('Institution', 'Reported_Gender'); #'FlowcellBatch'
#postAdjustCovars = c('Institution', 'Reported_Gender');
# Assign residual covariates
residualCovars = setdiff(preAdjustedSigCovars$significantCovars, c(postAdjustCovars, primaryVariable, 'Individual_ID'))
residualSigCovars = preAdjustedSigCovars
covariatesEffects = preAdjustedSigCovars$Effects.significantCovars[residualCovars]
postAdjustCovars = c(postAdjustCovars, names(which.max(covariatesEffects))) %>% unique()
notEstimable = c()

@@

  message( paste0(Tissue, ' Mean Correlatioon: ', mean(Correlation$Correlation)) )
  message( paste0(Tissue, ' Mean Std. Dev: ', sd(Correlation$Correlation)) )

  ggsave( filename=paste0( Tissue, "_Val.pdf" ),
    plot = ggplot( data = Correlation, aes( x = Correlation, y = Log10_Pval, col = Sig ) ) + 
      geom_point() + ggtitle("Log10( PVal) vs. Correlation" ) + 
      theme(plot.title = element_text(hjust = 0.5)) + ylab("-Log( PValue)") ,
    device = "pdf",
    path = NULL,
    scale = 1,
    width = 25,
    height = 8.75,
    units = "in",
    dpi = 300,
    limitsize = TRUE
  )

@@

#### Down Sample the DLPFC to look like ACC
```{r DownSample, cache=FALSE, fig.height=10, fig.width=8, results='asis'}
Tissue <- 'PFC'

Perm <- function( Exp, Impute ){  
  Template <- colnames(Exp)[ colnames(Exp) %in% colnames(Impute) ]
  SAMP <- sample(Template, 381, replace = FALSE, prob = NULL)
  Impute <- Impute[ , SAMP ]
  Exp <- Exp[ , SAMP ]

  Correlation <- as.data.frame(Correlation)
  Correlation$Sig <- 'NO'
  Correlation[ Correlation$FDR_W_Only < 0.05 & Correlation$Correlation > 0 , ]$Sig <- 'YES'
  Tab <- table(Correlation$Sig)
  return( t(as.data.frame(Tab))['Freq',] )
}

set.seed(42)
P <- 100
temp <- foreach( i=1:P ) %dopar% Perm( Exp, Impute )
Ps <- do.call( rbind, temp )
Ps <- as.data.frame( Ps )
Ps$V1 <- as.numeric(as.character(Ps$V1))
Ps$V2 <- as.numeric(as.character(Ps$V2))

MEAN <- mean( Ps[ , 2 ]/(  Ps[ , 2 ] +  Ps[ , 1 ]) )
print( paste0('Mean Downsampled Validation: ', MEAN ))
SD <- sd( Ps[ , 2 ]/(  Ps[ , 2 ] +  Ps[ , 1 ]))
error <- qt(0.975, df=dim(Ps)[1]-1)*sd( Ps[ , 2 ]/(  Ps[ , 2 ] +  Ps[ , 1 ]) )/sqrt( dim(Ps)[1] )

@@

#Set Used SynIDs For Provenance
Syns_Used <- c('syn21444814', 'syn21442621', 'syn16816490', 'syn16816488', 'syn21156161', 'syn16783596', 'syn16783867', 'syn2511399', 'syn16783557','syn16783508', 'syn16783557' )

@@

# Store SVA results
write.table(RESIDUAL.SVA.GENE_EXPRESSION, file = 'CMC_SVA_Diagnosis_Corrected.tsv', sep = '\t', quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='CMC_SVA_Diagnosis_Corrected.tsv', name = 'Normalised, covariates, surrogate and Diagnosis variable adjusted residual expression (for eQTL)', parentId=CODE$properties$id ), activityName = activityName, executed = c( Syns_Used, thisFile) , activityDescription = activityDescription)
  all.annotations$dataSubType = 'residualGeneExpForeQTL'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

# Store residual gene expression for network analysis
RESIDUAL.NET.GENE_EXPRESSION %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'CMC_Diagnosis_Regressed_netResidualExpression.tsv', sep = '\t', row.names=F, quote=F)
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='CMC_Diagnosis_Regressed_netResidualExpression.tsv', name = 'Normalised, covariates and Diagnosis removed residual expression (for network analysis)', parentId=CODE$properties$id ), activityName = activityName, executed = c( Syns_Used, thisFile) , activityDescription = activityDescription)
  all.annotations$dataSubType = 'residualGeneExpForNetAnlz'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

#Upload Plots:
ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ACC_Val.pdf', name = 'LogP value as a function of correlation for ACC', parentId=CODE$properties$id ), activityName = activityName, executed = c( Syns_Used, thisFile) , activityDescription = activityDescription)
  all.annotations$tissueTypeAbrv = 'ACC'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ACC_Cor_All.eps', name = 'Histogram of all correlation values for ACC', parentId=CODE$properties$id ), activityName = activityName, executed = c( Syns_Used, thisFile) , activityDescription = activityDescription)
  all.annotations$tissueTypeAbrv = 'ACC'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='ACC_Cor_gt5Perc.eps', name = 'Histogram of Significant correlation values for ACC', parentId=CODE$properties$id ), activityName = activityName, executed = c( Syns_Used, thisFile) , activityDescription = activityDescription)
  all.annotations$tissueTypeAbrv = 'ACC'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='PFC_Val.pdf', name = 'LogP value as a function of correlation for DLPFC', parentId=CODE$properties$id ), activityName = activityName, executed = c( Syns_Used, thisFile) , activityDescription = activityDescription)
  all.annotations$tissueTypeAbrv = 'DLPFC'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='PFC_Cor_All.eps', name = 'Histogram of all correlation values for DLPFC', parentId=CODE$properties$id ), activityName = activityName, executed = c( Syns_Used, thisFile) , activityDescription = activityDescription)
  all.annotations$tissueTypeAbrv = 'DLPFC'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)

ENRICH_OBJ <-  syn_temp$store( synapseclient$File( path='PFC_Cor_gt5Perc.eps', name = 'Histogram of Significant correlation values for DLPFC', parentId=CODE$properties$id ), activityName = activityName, executed = c( Syns_Used, thisFile) , activityDescription = activityDescription)
  all.annotations$tissueTypeAbrv = 'DLPFC'
  syn_temp$setAnnotations(ENRICH_OBJ, annotations = all.annotations)
